---
title: "GroupProject"
author: "Group3"
date: "11/23/2021"
output: pdf_document
---
## Loading Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
nh = data.frame(NHANES::NHANES)
library("Hmisc")
library("dplyr")
library("tidyr")
library("sas7bdat")
library("corrplot")
library("gtsummary")
library("dplyr")
library("olsrr")
library(kableExtra)
library(jtools)
```

## Data Cleaning and Variates Selection
```{r}
varinterst = c("Age","Gender","Race1","Depressed","Diabetes","BPSysAve", "BPDiaAve", "TotChol", "Testosterone", "RegularMarij", "Weight", "Height")
nh1 = nh[,varinterst]
nh1 <- nh1[ which(nh1$Age >=18 & nh1$Age <=59), ] #Due to Marijuana variable. 
nh1 <- nh1 %>%
  mutate(HypT = if_else(nh1$BPSysAve >= 130 | nh1$BPDiaAve >=80, 1, 0)) #Add hypertension variable
nh1 <- nh1 %>%
  mutate(Sex = ifelse(nh1$Gender == "female", 1, 0)) #Recode Sex into binary where female=1, male=0
nh1 <- nh1 %>%
  mutate(Diabete = ifelse(nh1$Diabetes == "No", 0, 1)) #Recode Diabetes into binary where No=0, Yes=1
nh1 <- nh1 %>%
  mutate(RegMarij = ifelse(nh1$RegularMarij == "No", 0, 1)) #Recode Regular Marij into binary where No=0, Yes=1
nh1 <- nh1 %>%
  mutate(Depress = ifelse(nh1$Depressed == "None", 0, 1)) #Recode Depressed into reports syptoms or does not have syptoms 
nh1$Incomplete = complete.cases(nh1) 
nh1$Race1 = as.factor(nh1$Race1) #Race as factor
#Creates Sex Specific Data Sets
drops = c("Sex")
nhF= nh1[ which(nh1$Sex == 1), ]
nhM = nh1[ which(nh1$Sex == 0), ]
nhF = nhF[,!(names(nhF) %in% drops)]
nhM = nhM[,!(names(nhM) %in% drops)]
```

## Complete/Incomplete Data Set
```{r}
nh1[,c("Age","Gender","Race1","Depressed","Diabetes","TotChol", "Testosterone", "RegularMarij", "Weight", "Height", "HypT", "Incomplete")] %>%
 tbl_summary(by = Incomplete,
              missing = "no",
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              ) %>%
  add_n %>%
  add_p(test = list(Age ~ "t.test",
                    Race1 ~ "chisq.test",
                    Depressed ~ "chisq.test",
                    Diabetes ~ "chisq.test",
                    HypT ~ "chisq.test",
                    TotChol ~ "t.test",
                    Testosterone ~ "t.test",
                    RegularMarij ~ "chisq.test",
                    Weight ~ "t.test",
                    Height ~ "t.test"),
        test.args = all_tests("t.test") ~ list(var.equal = TRUE),## Important argument! 
        pvalue_fun = function(x) style_pvalue(x, digits = 2)) %>%
  bold_p(t = 0.05) %>%
  bold_labels
```
## Female only Complete/Incomplete Data Set
```{r}
## Shiny app
nhF[,c("Age","Gender","Race1","Depressed","Diabetes","TotChol", "Testosterone", "RegularMarij", "Weight", "Height", "HypT", "Incomplete")] %>%
 tbl_summary(by = Incomplete,
              missing = "no",
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              ) %>%
  add_n %>%
  add_p(test = list(Age ~ "t.test",
                    Race1 ~ "chisq.test",
                    Depressed ~ "chisq.test",
                    Diabetes ~ "chisq.test",
                    HypT ~ "chisq.test",
                    TotChol ~ "t.test",
                    Testosterone ~ "t.test",
                    RegularMarij ~ "chisq.test",
                    Weight ~ "t.test",
                    Height ~ "t.test"),
        test.args = all_tests("t.test") ~ list(var.equal = TRUE),## Important argument! 
        pvalue_fun = function(x) style_pvalue(x, digits = 2)) %>%
  bold_p(t = 0.05) %>%
  bold_labels
```

##Exploratory Stats: Male only Complete/Incomplete Data Set
```{r}
nhM[,c("Age","Gender","Race1","Depressed","Diabetes","TotChol", "Testosterone", "RegularMarij", "Weight", "Height", "HypT", "Incomplete")] %>%
 tbl_summary(by = Incomplete,
              missing = "no",
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              ) %>%
  add_n %>%
  add_p(test = list(Age ~ "t.test",
                    Race1 ~ "chisq.test",
                    Depressed ~ "chisq.test",
                    Diabetes ~ "chisq.test",
                    HypT ~ "chisq.test",
                    TotChol ~ "t.test",
                    Testosterone ~ "t.test",
                    RegularMarij ~ "chisq.test",
                    Weight ~ "t.test",
                    Height ~ "t.test"),
        test.args = all_tests("t.test") ~ list(var.equal = TRUE),## Important argument! 
        pvalue_fun = function(x) style_pvalue(x, digits = 2)) %>%
  bold_p(t = 0.05) 
```

Our first limitation is that complete/incomplete data are significantly different. Even if we adjusted for gender, there is still significant difference between the two data set. Using the complete model can be biased. Not generalizable to US population. 

```{r}
nh2 = drop_na(nh1) #complete data set
```

## Exploratory Stats: Testosterone (and other continous)  by Categorical Variables Tables (assess) (NOT INCLUDE IN PAPER OR PRESENTATION DIRECTLY )
```{r, echo=FALSE, eval=FALSE}
SummaryTestTable <- function(x){ 
nh3 = nh2[,c(all_of(x),"Testosterone", "Age", "TotChol", "Weight", "Height")]
  nh3 %>%
   tbl_summary(by = all_of(x),
                missing = "no",
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              ) %>%
  add_n %>%
  add_p(test = list(Testosterone ~ "t.test",
                    Age ~ "t.test",
                    TotChol ~ "t.test",
                    Weight ~ "t.test",
                    Height ~ "t.test"),
        test.args = all_tests("t.test") ~ list(var.equal = TRUE),## Important argument! 
        pvalue_fun = function(x) style_pvalue(x, digits = 2)) %>%
  bold_p(t = 0.05) %>%
  bold_labels}
for (i in c("Gender","Diabetes", "HypT", "RegularMarij", "Depress")) {
  print(SummaryTestTable(i))
}
nh3 = nh2[,c("Race1","Testosterone", "Age", "TotChol", "Weight", "Height")]
  nh3 %>%
   tbl_summary(by = Race1,
                missing = "no",
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              ) %>%
  add_n %>%
 add_p(test = list(Testosterone ~ "aov",
                    Age ~ "aov",
                    TotChol ~ "aov",
                    Weight ~ "aov",
                    Height ~ "aov"),
        pvalue_fun = function(x) style_pvalue(x, digits = 2)) %>%
  bold_p(t = 0.05) %>%
  bold_labels
  
```
## Exploratory Stats: Continous Variables (NOT INCLUDE IN PAPER OR PRESENTATION DIRECTLY )
```{r}
#Compare Y to continuous X
scatmatrixData = nh2[,c("Testosterone", "TotChol", "Age", "Height", "Weight")]
panel.hist <- function(x, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}
pairs(scatmatrixData, pch = 19, diag.panel=panel.hist)
cormat = cor(scatmatrixData)
pres <- cor.mtest(scatmatrixData, conf.level = .95)
corrplot.mixed(cormat, lower.col = "black", number.cex = 1,p.mat = pres$p, sig.level = .05)
```
## Now center continous variables (Needed)
```{r}
nh2$Weightc = (nh2$Weight-mean(nh2$Weight, na.rm = TRUE))/sd(nh2$Weight, na.rm = TRUE)
nh2$Heightc = (nh2$Height-mean(nh2$Height, na.rm = TRUE))/sd(nh2$Height, na.rm = TRUE)
nh2$Agec = (nh2$Age-mean(nh2$Age, na.rm = TRUE))/sd(nh2$Age, na.rm = TRUE)
nh2$TotCholc = (nh2$TotChol-mean(nh2$TotChol, na.rm = TRUE))/sd(nh2$TotChol, na.rm = TRUE)
```
## Function to check assumptions 
```{r}
checka <- function (model) {
  car::avPlots(model)
  plot(model$fitted.values, rstudent(model))
  hist(rstudent(model))
  car::qqPlot(rstudent(model))
  shapiro.test(rstudent(model)
               ) ##Should we include shapiro wilk????
}
``` 
## Unrefined Model (what to say about this?)
```{r}
m_unrefined <- lm(Testosterone~RegMarij, data=nh2)
summary(m_unrefined)
checka(m_unrefined) ##Comment on plots
```

## Adjusted Model with no log 
```{r}
m_adjusted_nolog =  lm(Testosterone~ RegMarij + Sex + Agec + Race1  + Depress + Weightc + Heightc +
             Diabete + HypT + TotCholc, 
         data = nh2)
summary(m_adjusted_nolog)
checka(m_adjusted_nolog) ##Constant variance. We think the two clusters are due to sex. Also trend within a cluster. 
```
## Both models above had linear trend in plots, so we adjusted model log(Y)-THIS IS FINAL MAIN EFFECTS MODEL
```{r}
m_adjust <- lm(log(Testosterone)~ RegMarij + Sex + Agec + Race1  + Depress + Weightc + Heightc +
             Diabete + HypT + TotCholc, data = nh2)
summary(m_adjust)
checka(m_adjust)
```

## Due to constant variance plot having clusters and literature, we think that TT distribution by sex underlies the trend. Test this by stratifying model. See that in stratified models, this is elimnated 
```{r}
nhFemale = nh2[ which(nh2$Sex == 1), ]
nhMale = nh2[ which(nh2$Sex == 0), ]
m_F = lm(log(Testosterone)~ RegMarij + Agec + Race1  + Depress + Weightc + Heightc +
             Diabete + HypT + TotCholc,
         data = nhFemale)
summ(m_F)
plot(m_F$fitted.values, m_F$residuals)
m_M = lm(log(Testosterone)~ RegMarij + Agec + Race1  + Depress + Weightc + Heightc +
             Diabete + HypT + TotCholc,
         data = nhMale)
summ(m_M)
plot(m_F$fitted.values, m_F$residuals)
```

## Multicolinearity check
```{r}
car::vif(m_adjusted_nolog)
```







##Adding interactions. Based on analysis and stratified models, we assume Sex has effect modifcation on variables. 
```{r}
m_int <- lm(lm(log(Testosterone)~ RegMarij + Sex + Agec + Race1  + Depress + Weightc + Heightc +
             Diabete + HypT + TotCholc + Sex*(RegMarij + Agec + Race1  + Depress + Weightc + Heightc +
             Diabete + HypT + TotCholc),  data = nh2))
summary(m_int)

```
## Remove all insig interactions, and checks assumptions (not linearity, as this is checked only for main effects (see above))
```{r}
m_final <- lm(lm(log(Testosterone)~ RegMarij + Sex + Agec + Race1  + Depress + Weightc + Heightc +
             Diabete + HypT + TotCholc + Sex*(Agec + Race1  + RegMarij +  Weightc + Heightc +
             Diabete + TotCholc),  data = nh2))
summary(m_final)
plot(m_final$fitted.values, m_final$residuals)
hist(m_final$residuals)
car::qqPlot(m_final$residuals, main = "QQ plot", xlab = "Normal Quantiles", 
            ylab = "Residuals")
shapiro.test(m_final$residuals)
summ(m_final)
```
```{r}
olsrr::ols_plot_dffits(m_final)
olsrr::ols_plot_dfbetas(m_final)
olsrr::ols_plot_cooksd_chart(m_final)
m_alt_cov = covratio(m_final)
n = nrow(nh2) 
p = m_final$rank
plot(m_alt_cov); abline(1+3*p/n,0); abline(1-3*p/n,0)
```


## For discussion, in main effects model, test if mean test is same for race groups (not something we need to include, but idk, we can get rid of, just thought it was a way to utilize GLH learned in class)
```{r}
m_wo_its =  lm(log(Testosterone)~ -1 + RegMarij + Sex + Agec + Race1  + Depress + Weightc + Heightc +
             Diabete + HypT + TotCholc, data = nh2)
summary(m_wo_its)
#Test if all five are same
Contrast.T = matrix(c(0,0,0,1,-1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,-1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,-1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,-1,0,0,0,0,0,0),byrow = T, nrow = 4) 
car::linearHypothesis(model=m_wo_its,hypothesis.matrix=Contrast.T, rsh=c(0,0,0,0))
#Test if (4) are same 
Contrast.T = matrix(c(0,0,0,0,1,-1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,-1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,-1,0,0,0,0,0,0),byrow = T, nrow = 3)
car::linearHypothesis(model=m_wo_its,hypothesis.matrix=Contrast.T, rsh=c(0,0,0))
```

## Additional Analysis (Leah)

```{r}
summary(nh2$Testosterone)
hist(nh2$Testosterone)
hist(log(nh2$Testosterone), main = "Histogram of log(Testosterone)", 
     xlab = "log(Testosterone) (log(ng/dL))", col = "grey")
```

```{r}
m_int <- lm(lm(log(Testosterone)~ RegMarij + Sex + Agec + Race1  
               + Depress + Weightc + Heightc + Diabete + HypT + 
                 TotCholc + Sex*RegMarij,  data = nh2))
summary(m_int)
```


```{r}
md = lm(formula = log(Testosterone) ~ RegMarij + Sex + Agec + Race1 + 
    Depress + Weightc + Heightc + Diabete + HypT + TotCholc, 
    data = nh2)
md$col = ifelse(nh2$Sex == 1, 'red', 'blue')
plot(md$fitted.values, rstudent(md), col = md$col, 
     main = "Equal Variance Plot for main effect model (with log transform)", 
     xlab = "Fitted Values", ylab = "Studentized Residuals")
legend("topright", legend = c("female", "male"), col = c('red','blue'), 
       pch = 1)
```


```{r}
md = lm(formula = Testosterone ~ RegMarij + Sex + Agec + Race1 + 
    Depress + Weightc + Heightc + Diabete + HypT + TotCholc, 
    data = nh2)
md$col = ifelse(nh2$Sex == 1, 'red', 'blue')
plot(md$fitted.values, rstudent(md), col = md$col, 
     main = "Equal Variance Plot for main effect model (w/o log transform)")
legend("topright", legend = c("female", "male"), col = c('red','blue'), 
       pch = 1)
```
